---
title: "Lecture 10 Code Examples"
output: 
  pdf_document:
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(fpp3)
```

# Misspecified ARMA models

Let's generate data from an AR(2) model. What happens if we fit a misspecified 
model, i.e. if we choose the wrong order for $p$.

```{r create-ar-data-and-fit-models}
set.seed(5209)
ar2_data <- arima.sim(model = list(ar = c(0.6, -0.2)), n = 100)
ar1_model <- arima(ar2_data, order = c(1, 0, 0))
ar2_model <- arima(ar2_data, order = c(2, 0, 0))
ar4_model <- arima(ar2_data, order = c(4, 0, 0))
```

```{r repeat-experiment-several-times}
set.seed(5209)
B <- 500
ar2_data <- map(1:B, ~ arima.sim(model = list(ar = c(0.6, -0.2)), n = 100))
ar1_model_coefs_ <- map(ar2_data, ~ arima(., order = c(1, 0, 0))$coef) |>
  transpose() |>
  map(unlist) |>
  as.tibble()
ar2_model_coefs_ <- map(ar2_data, ~ arima(., order = c(2, 0, 0))$coef) |>
  transpose() |>
  map(unlist) |>
  as.tibble()
ar4_model_coefs_ <- map(ar2_data, ~ arima(., order = c(4, 0, 0))$coef) |>
  transpose() |>
  map(unlist) |>
  as.tibble()
```

```{r plot-phi1-coefs}

phi1_coefs <- tibble(ar1 = ar1_model_coefs_$ar1,
                     ar2 = ar2_model_coefs_$ar1,
                     ar4 = ar4_model_coefs_$ar1)
phi1_coefs |> 
  pivot_longer(cols = everything(),
               names_to = "model",
               values_to = "coefs") |>
  ggplot() +
  geom_density(aes(x = coefs, color = model)) + 
  geom_vline(xintercept = 0.6)
```
```{r plot-phi2-coefs}

phi2_coefs <- tibble(ar2 = ar2_model_coefs_$ar2,
                     ar4 = ar4_model_coefs_$ar2)
phi2_coefs |> 
  pivot_longer(cols = everything(),
               names_to = "model",
               values_to = "coefs") |>
  ggplot() +
  geom_density(aes(x = coefs, color = model)) + 
  geom_vline(xintercept = -0.2)
```

# Real data analysis

## Working with tsibble

The `tsibble` package allows us to work with multiple time series in one data frame.
For instance, consider the `global_economy` data frame, which contains economic 
indicators featured by the World Bank from 1960 to 2017.
Each time series is identified by a `Key`.
The time series may be multivariate, i.e. have multiple columns.

## CAF exports

Let us try to model exports from the Central African Republic.

```{r}
caf_economy <- global_economy |>
  filter(Code == "CAF")
caf_economy |>
  autoplot(Exports) +
  labs(title="Central African Republic exports",
       y="% of GDP")
```

This is non-stationary, so we can take a first difference.


```{r}
caf_economy |>
  gg_tsdisplay(difference(Exports), plot_type='partial')
```

From the ACF and PACF plots, it seems that either an AR(2) or MA(3) model is 
appropriate for the residuals. We hence fit these two models, and also try
automatic model search (we will discuss this more next week). We also fit an AR(5)
model for comparison. The `fable` package makes fitting all 3 models at the 
same time extremely easy. The result is a `mable`, i.e. a dataframe of models.

```{r}
caf_fit <- caf_economy |>
  model(arima210 = ARIMA(Exports ~ pdq(2,1,0)),
        arima013 = ARIMA(Exports ~ pdq(0,1,3)),
        arima510 = ARIMA(Exports ~ pdq(5,1,0)),
        auto = ARIMA(Exports))

caf_fit |> glance()
```

We see that while ARIMA(5,1,0) has the largest log likelihood, it has the largest
AIC and AICc (smaller is better). The AIC and AICc of the other 3 models are
comparable. Finally, we check the order of the model found by automatic model search:
We got an ARIMA(2,1,2) model.

```{r}
caf_fit["auto"]
```

We now do a residual diagnosis. To see if 

```{r}
caf_fit |>
  select(arima210) |>
  gg_tsresiduals()
```

The `augment` method produces the fitted and residual values for each model.

```{r}
augment(caf_fit)
```

We can now use the residuals to compute a Ljung-Box test statistic for each model. 
We see that the p-values are large, so in each case, the residuals are 
well-approximated by a white noise sequence.

```{r}
augment(caf_fit) |>
  # filter(.model=='arima210') |>
  features(.innov, ljung_box, lag = 10, dof = 3)
```

Finally, we can forecast using our model.

```{r}
caf_fit |>
  forecast(h=5) |>
  filter(.model=='arima210') |>
  autoplot(global_economy)
```
## CAF GDP: Understanding ARIMA models

```{r}
caf_economy |>
  autoplot(GDP) +
  labs(title="Central African Republic GDP")
```

```{r}
caf_economy |>
  gg_tsdisplay(difference(GDP), plot_type='partial')
```
```{r}
caf_fit <- caf_economy |>
  model(arima200 = ARIMA(GDP ~ pdq(2,0,0)),
        arima210 = ARIMA(GDP ~ pdq(2,1,0)),
        arima220 = ARIMA(GDP ~ pdq(2,2,0)),
        auto = ARIMA(GDP))

caf_fit |> glance()
```
```{r}
caf_fit |>
  forecast(h=5) |>
  autoplot(global_economy)
```

